### 2015-03-28 drafts/anderson-etal-fisheries-ensembles.Rmd
Add some notes

---
output: html_document
bibliography: ../notes/fisheries-ensembles.bib
csl: ../notes/cjfas.csl
---

# Combining data-limited fisheries models to derive robust estimates of stock status

Sean C. Anderson, Jamie Afflerbach, Andrew B. Cooper, Mark Dickey-Collas, Olaf P. Jensen, Kristin M. Kleisner, Catherine Longo, Cóilín Minto, Giacomo Chato Osio, Dan Ovando, Andrew A. Rosenberg, Elizabeth R. Selig, James T. Thorson, and potentially others...

# Introduction

@greene2006: probably the closest thing to what we're doing but with climate models (Bayesian multilevel ensemble)

# Methods

# Results

# Discussion

# Acknowledgements

# Figures

# References