### 2015-06-09 analysis/plot-sims.tex
Add plots comparing differences between simulation scenarios

\documentclass[12pt]{article}
\usepackage{geometry}
\geometry{verbose, letterpaper, tmargin = 2.54cm, bmargin = 2.54cm,
  lmargin = 2.54cm, rmargin = 2.54cm}
\geometry{letterpaper}
\usepackage{graphicx}
\begin{document}

\begin{figure}[htbp]
\begin{center}
  \includegraphics[width=\textwidth]{../figs/sim-catch-by-sigmaC-iter1.pdf}
\caption{Time series of catch split by measurement error on catch. (All
  figures are for iteration 1 and demersal life history.) Seems like a
  judgement call on whether these are too similar. Hard to say. See the next
  plots that explore how these affect the model estimates.}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
  \includegraphics[width=\textwidth]{../figs/sim-CMSY-by-sigmaC-iter1.pdf}
\caption{Time series of CMSY B/BMSY split by measurement error on catch.
  These, for the most part, are pretty similar.}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
  \includegraphics[width=\textwidth]{../figs/sim-COMSIR-by-sigmaC-iter1.pdf}
\caption{Time series of COMSIR B/BMSY split by measurement error on catch.
  Again, fairly similar, but not sure if too similar.}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
  \includegraphics[width=\textwidth]{../figs/sim-SSCOM-by-sigmaC-iter1.pdf}
\caption{Time series of SSCOM B/BMSY split by measurement error on catch.
  SSCOM is more variable here than I thought it would be.}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
  \includegraphics[width=\textwidth]{../figs/sim-catch-by-AR-iter1.pdf}
\caption{Time series of catch split by recruitment deviation autocorrelation.
These look reasonably different in most cases.}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
  \includegraphics[width=\textwidth]{../figs/sim-catch-by-TS-iter1.pdf}
\caption{Time series of catch split by time series length. I.e., I
    thought these should match exactly. Not sure what's happening here,
    although my guess is that one or more of the stochastic draws is based on
  a different seed.}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
  \includegraphics[width=\textwidth]{../figs/sim-CMSY-by-TS-iter1.pdf}
\caption{Time series of CMSY B/BMSY split by time series length. These seem to
be deviating pretty strongly from each other, as we might expect having played
with these models before.}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
  \includegraphics[width=\textwidth]{../figs/sim-catch-by-sigmaR-iter1.pdf}
\caption{Time series of catch split by magnitude of recruitment variability.
  Again, tough call.}
\end{center}
\end{figure}

\end{document}