### 2017-07-18 analysis/99-ohi.Rmd
Add OHI ROC analysis

---
title: "ROC for OHI thresholds"
author: "Sean Anderson"
date: '2017-07-18'
output: html_document
---

Let's read in the data, source some functions, and manipulate the data so we can create some ROC curves at relevant thresholds for the OHI.

```{r}
library(dplyr)
library(ggplot2)
source("0-ensemble-functions.R")

# bring in the simulated data and shape it for plotting:
clean_names <- dplyr::data_frame(
  method = c("CMSY", "COMSIR", "mPRM", "SSCOM",
    "gbm_ensemble", "rf_ensemble", "lm_ensemble", "mean_ensemble"),
  clean_method = c("CMSY", "COM-SIR", "mPRM", "SSCOM",
    "GBM Superensemble", "RF Superensemble", "LM Superensemble", "Mean Ensemble"),
  order = c(1, 2, 4, 3, 8, 7, 6, 5),
  label_fudge_x = c(
    0, -0.03,0,-0.04,
    0, 0,0,-0.05),
  label_fudge_y = c(
    0, -0.02,0.03,0.04,
    0, 0.02,0,-0.02))

d_sim <- readRDS("generated-data/cv_sim_long.rds")
d_sim <- suppressWarnings(inner_join(d_sim, clean_names))

get_performance <- function(validation_data) {
  d_sim_perf <- validation_data %>% group_by(test_iter, type, clean_method,
    order, label_fudge_x, label_fudge_y) %>%
    summarise(
      mare = median(abs(re)),
      mre = median(re),
      corr = cor(bbmsy_true_trans, bbmsy_est_trans, method = "spearman",
        use = "pairwise.complete.obs"))

  d_sim_perf$clean_method <- as.factor(d_sim_perf$clean_method)
  d_sim_perf$clean_method <- reorder(d_sim_perf$clean_method, d_sim_perf$order)

  d_sim_perf_slope <- filter(d_sim_perf, type == "slope")
  d_sim_perf <- filter(d_sim_perf, type == "mean")

  d_sim_perf_long <- reshape2::melt(d_sim_perf,
    id.vars = c("test_iter", "clean_method", "type", "order", 
      "label_fudge_x", "label_fudge_y"))
  d_sim_perf_summ <- d_sim_perf_long %>%
    group_by(clean_method, order, variable, label_fudge_x, label_fudge_y) %>%
    summarise(
      m = median(value),
      l = quantile(value, 0.25),
      u = quantile(value, 0.75)) %>%
    as.data.frame
  d_sim_perf_summ$text_col <- ifelse(grepl("ensemble", 
    d_sim_perf_summ$clean_method, ignore.case = TRUE),
    "grey20", "grey50")
  d_sim_perf_wide_temp <- reshape2::dcast(d_sim_perf_summ, 
    clean_method + order ~ variable,
    value.var = "m")
  d_sim_perf_wide_temp
}
d_sim_perf_wide <- get_performance(d_sim)
saveRDS(d_sim_perf_wide, "generated-data/d_sim_perf_wide.rds")

# Format basic simulation ensemble without covariates for plotting:
d_sim_basic <- readRDS("generated-data/cv_sim_mean_basic_long.rds")
d_sim_basic <- suppressWarnings(inner_join(d_sim_basic, clean_names))
d_sim_perf_wide_basic <- get_performance(d_sim_basic)
saveRDS(d_sim_perf_wide_basic, "generated-data/d_sim_perf_wide_basic.rds")
d_sim_basic$bbmsy_est[d_sim_basic$bbmsy_est > 10] <- NA
d_sim_basic <- na.omit(d_sim_basic)
```

Get ROC and AUC for models applied to simulation data with b/bmsy as the response. We'll use 0.8 B/Bmsy as the threshold to start with.

```{r}
library("doParallel")
registerDoParallel(cores = parallel::detectCores())
.parallel <- ifelse(Sys.info()[["sysname"]] == "Windows", FALSE, TRUE)

d_sim_mean <- d_sim_basic %>% filter(type == "mean") %>%
  mutate(above_0.8 = ifelse(bbmsy_true > 0.8, 1, 0)) %>% 
  mutate(above_1.45 = ifelse(bbmsy_true > 1.45, 1, 0)) %>% 
  mutate(below_1.45_above_0.8 = ifelse(bbmsy_true < 1.45 & bbmsy_true > 0.8, 1, 0)) %>% 
  mutate(clean_method = paste(order, clean_method, sep = "-"))

d_sim_mean$Ensemble <- ifelse(grepl("ensemble", 
  d_sim_mean$clean_method, ignore.case = TRUE),
  "(a) Ensemble", "(b) Individual")

get_roc <- function(true, est) {
  library("pROC") # required or pROC:: will generate errors
  y <- pROC::roc(response = true, predictor = est)
  data.frame(sens = y$sensitivities, spec = y$specificities,
    auc = as.numeric(y$auc))
}

rocs_sim_mean_0.8 <- plyr::ddply(d_sim_mean, c("clean_method", "Ensemble"), 
  function(x) {
    get_roc(true = x$above_0.8, est = x$bbmsy_est)
  }, .parallel = .parallel)

rocs_sim_mean_1.45 <- plyr::ddply(d_sim_mean, c("clean_method", "Ensemble"), 
  function(x) {
    get_roc(true = x$above_1.45, est = x$bbmsy_est)
  }, .parallel = .parallel)

rocs_sim_mean_0.8_1.45 <- plyr::ddply(d_sim_mean, c("clean_method", "Ensemble"), 
  function(x) {
    get_roc(true = x$below_1.45_above_0.8, est = x$bbmsy_est)
  }, .parallel = .parallel)

saveRDS(rocs_sim_mean_0.8, file = "generated-data/rocs_sim_mean_0.8.rds")
saveRDS(rocs_sim_mean_1.45, file = "generated-data/rocs_sim_mean_1.45.rds")
saveRDS(rocs_sim_mean_0.8_1.45, file = "generated-data/rocs_sim_mean_0.8_1.45.rds")

# auc_sim_mean <- rocs_sim_mean %>%
#   group_by(clean_method) %>%
#   summarise(auc = auc[1])
```

Now make the curves:

```{r}
make_roc_curves <- function(dat) {
  p <- ggplot(dat, aes(spec, sens, colour = clean_method, group = clean_method)) +
    geom_line() + coord_equal() +
    geom_abline(intercept = 1, slope = 1, lty = 2, col = "darkgrey") +
    xlim(1, 0) +
    xlab(expression(Rate~of~correctly~categorizing~B/B[MSY]~as < threshold)) +
    ylab(expression(Rate~of~correctly~categorizing~B/B[MSY]~as > threshold)) +
    theme_bw() +
    theme(plot.background = element_blank(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()) + facet_wrap(~Ensemble) +
    scale_colour_brewer(palette="Spectral", guide = guide_legend(title = "Model"))
  p
}
p <- make_roc_curves(rocs_sim_mean_0.8)
ggsave("../figs/ohi-roc-sim-0.8.png", width = 8, height = 5)

p <- make_roc_curves(rocs_sim_mean_1.45)
ggsave("../figs/ohi-roc-sim-1.45.png", width = 8, height = 5)

p <- make_roc_curves(rocs_sim_mean_0.8_1.45)
ggsave("../figs/ohi-roc-sim-0.8-1.45.png", width = 8, height = 5)
```